Predicting Kamala Harris's Victory in the 2024 US Election

### 2024-10-31 paper/paper.qmd
Add the Historical Lean of the Missing State

The dataset for this analysis is sourced from FiveThirtyEight, which rigorously collects and aggregates polling data from a variety of firms to reflect public opinion. To ensure the integrity of the data, only polls that adhere to specific criteria are included in the dataset. Each poll must provide essential information, such as the pollster’s name, survey dates, sample sizes, and methodological details (e.g., polling medium, voter files, weighting criteria). Polls that are deemed nonscientific, that blend data from multiple sources, or that are conducted by hobbyists are excluded.

Once a poll meets these stringent standards, it is incorporated into the database, enabling it to inform polling averages, forecasts, and political coverage. This careful selection process ensures that the dataset accurately captures and reflects the nuances of public sentiment and behavior.

@@

# Define electoral votes for each state
electoral_votes <- data.frame(
  state = c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida",
            "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", 
            "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", 
            "Missouri", "Montana", "Nebraska", "Nevada", 
            "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", 
            "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", 
            "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming", 
            "District of Columbia"),
  electoral_votes = c(9, 3, 11, 6, 55, 9, 7, 3, 29, 16, 4, 4, 20, 11, 6, 6, 8, 8, 4, 10, 11, 15, 10, 6, 10, 
                      3, 10, 6, 4, 14, 5, 29, 16, 3, 18, 7, 6, 20, 4, 9, 3, 11, 38, 6, 3, 13, 12, 5, 10, 3, 3))

@@

# Print the results for the total electoral votes
print(paste("Trump Electoral Votes:", trump_electoral_votes))
print(paste("Harris Electoral Votes:", harris_electoral_votes))
```

@@

```{r tbl-prediction}
#| echno: false
#| label: fig-statemap
#| tbl-cap: Predicted Winner of 2024 US Presidential Election by State
#| echo: false
#| warning: false
#| message: false

# Replace NA values in the 'winner' column with 'Trump'
#state_predictions$winner[is.na(state_predictions$winner)] <- "Trump"

# "red" for Trump and "blue" for Harris
state_predictions$winner_color <- ifelse(state_predictions$winner == "Trump", "red", "blue")

@@

>>>>

The dataset used in this analysis originates from FiveThirtyEight, a trusted source known for its rigorous methodology in collecting and aggregating polling data to represent public opinion accurately. To maintain data integrity and ensure a high standard of reliability, only polls meeting stringent quality criteria are included in the dataset. Specifically, each poll must disclose crucial information, including the pollster’s name, the exact dates of the survey, the sample size, and the full methodological details. These details encompass the polling medium (e.g., phone, online), use of verified voter files, demographic weighting criteria, and any adjustments applied to reflect a representative sample.

Polls are rigorously verified, and those that do not meet these requirements are excluded from the dataset. For instance, polls categorized as nonscientific, those that merge data from multiple unrelated sources, and those conducted by hobbyists or unverified entities are filtered out to prevent data quality compromise.

Once deemed credible, polls are incorporated into the FiveThirtyEight database. These polls then contribute to polling averages, forecast models, and political coverage, which depend heavily on the quality and reliability of public opinion data. This meticulous selection process ensures that the dataset captures a comprehensive and nuanced view of public sentiment, providing an accurate reflection of behavioral trends and preferences over time.

@@

# Define electoral votes for each state
electoral_votes <- data.frame(
  state = c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida",
            "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", 
            "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", 
            "Missouri", "Montana", "Nebraska", "Nevada", 
            "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", 
            "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", 
            "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming", 
            "District of Columbia"),
  electoral_votes = c(9, 3, 11, 6, 54, 10, 7, 3, 30, 16, 4, 4, 20, 11, 6, 6, 8, 8, 4, 10, 11, 15, 10, 6, 10, 
                      4, 5, 6, 4, 14, 5, 28, 16, 3, 17, 7, 6, 19, 4, 9, 3, 11, 40, 6, 3, 13, 12, 5, 10, 3, 3))

@@

```{r}
#| echno: false
#| label: tbl-missing
#| tbl-cap: Missing States with Predicted Winner Based on Historical Lean
#| echo: false
#| warning: false
#| message: false
# Create the data frame with missing states, electoral votes, and predicted winner based on historical lean
missing_states <- data.frame(
  state = c("Alabama", "Alaska", "Arkansas", "Connecticut", "Delaware", "Hawaii", "Idaho", 
            "Illinois", "Indiana", "Kansas", "Kentucky", "Louisiana", "Mississippi", "North Dakota", 
            "Oklahoma", "Oregon", "South Carolina", "Tennessee", "Utah", "Washington", 
            "West Virginia", "Wyoming", "District of Columbia"),
  Electoral_Votes = c(9, 3, 6, 7, 3, 4, 4, 19, 11, 6, 8, 8, 6, 3, 7, 8, 9, 11, 6, 12, 4, 3, 3),
  Prediction_Based_on_Historical_Lean = c("Trump", "Trump", "Trump", "Harris", "Harris", 
                                          "Harris", "Trump", "Harris", "Trump", "Trump", 
                                          "Trump", "Trump", "Trump", "Trump", "Trump", 
                                          "Harris", "Trump", "Trump", "Trump", "Harris", 
                                          "Trump", "Trump", "Harris")
)

# Display the table with kable
kable(missing_states, col.names = c("State", "Electoral Votes", "Prediction Based on Historical Lean"))

@@

```{r tbl-prediction}
#| echno: false
#| label: fig-statemap
#| tbl-cap: Predicted Winner of 2024 US Presidential Election by State
#| echo: false
#| warning: false
#| message: false
state_predictions <- merge(state_predictions, missing_states, by = "state", all = TRUE)

state_predictions <- state_predictions %>%
  mutate(
    trump_electoral_votes = case_when(
      is.na(Prediction_Based_on_Historical_Lean) ~ trump_electoral_votes,  # Keep original if NA
      Prediction_Based_on_Historical_Lean == "Trump" ~ Electoral_Votes,    # Assign Electoral_Votes if "Trump"
      TRUE ~ 0  # Set to 0 otherwise
    ),
    harris_electoral_votes = case_when(
      is.na(Prediction_Based_on_Historical_Lean) ~ harris_electoral_votes,  # Keep original if NA
      Prediction_Based_on_Historical_Lean == "Harris" ~ Electoral_Votes,    # Assign Electoral_Votes if "Harris"
      TRUE ~ 0  # Set to 0 otherwise
    )
  )

# Replace NA values in the 'winner' column with historical lean predictions for missing states
state_predictions$winner <- ifelse(
  is.na(state_predictions$winner),
  state_predictions$Prediction_Based_on_Historical_Lean,
  state_predictions$winner
)

# Assign colors based on the updated 'winner' column
state_predictions$winner_color <- ifelse(state_predictions$winner == "Trump", "red", "blue")

@@

```{r}
#| echno: false
#| label: fig-statenum
#| tbl-cap: Numerical Predicted Winner of 2024 US Presidential Election by State
#| echo: false
#| warning: false
#| message: false

trump_total <- sum(state_predictions$trump_electoral_votes, na.rm = TRUE)
harris_total <- sum(state_predictions$harris_electoral_votes, na.rm = TRUE)

# Create a summary data frame
summary_df <- data.frame(
  Candidate = c("Trump", "Harris"),
  Total_Electoral_Votes = c(trump_total, harris_total)
)

# Display the summary table in kable format
cat("\nTotal Electoral Votes:\n")
kable(summary_df, col.names = c("Candidate", "Total Electoral Votes"), format = "markdown")
```
